---
title: "Business Entities in Larimer County"
author: "JBaker"
date: "March 7, 2018"
output: html_document
---
## Introduction
The following is some analysis of businesses in Larimer County. All source data was obtained from a download from http://data.colorado.gov  The data set contains historical data and is quite large (1.7M records). The data set description can be found here: https://data.colorado.gov/Business/Business-Entities-in-Colorado/4ykn-tg5h 

The basic workflow is: 

* Pare down the data set to Larimer County only.
* Select the years of interest
* Summarize the business opens and closes for the year
* Visualize the data

### Data Intake

```{r initialize, cache=TRUE}
# Download the data using: https://data.colorado.gov/api/views/4ykn-tg5h/rows.csv?accessType=DOWNLOAD&api_foundry=true

colclass <- c("numeric", "character", "character", "character", "factor", "factor", "character", 
              "factor", "character", "character", "factor", "factor", "character", "factor",
              "factor", "factor", "factor", "factor", "factor", "factor", "factor", "factor",
              "factor", "factor", "factor", "factor", "factor", "factor", "factor", "factor",
              "factor", "factor", "factor", "factor", "character") 

CObiz <- read.csv("./Business_Entities_in_Colorado.csv", colClasses = colclass)

LCzips <- c(80511, 80512, 80513, 80515, 80517, 80521, 80522, 80524, 80525, 80526, 80527, 80528, 80528, 80532, 80535, 80536, 80536, 80537, 80538, 80539, 80541, 80545, 80547, 80549)

```

## Transform and Cull the Data
Select out the entries for zip codes corresponding to Larimer County.

```{r cull, warning=FALSE}
# Replace zipcodes of form #####-#### with just #####

CObiz$principalzipcode <- as.numeric(sub("([0-9]+)-([0-9]+)", "\\1", CObiz$principalzipcode))
LCbiz <- CObiz[which(CObiz$principalzipcode %in% LCzips), ]
LCgoodstanding <- nrow(LCbiz[LCbiz$entitystatus == "Good Standing", ])
```

There are currently `r LCgoodstanding` Larimer County businesses with a "Good Standing" status.

## Summarize the Entity Formation and Dissolution by Year

```{r summarize, message=FALSE}
LCsumm <- NULL
for (i in seq(2000, 2017)) {
        dissolve <- length(grep(as.character(i), LCbiz$entityname))
        init <- length(grep(as.character(i), LCbiz$entityformdate))
        LCsumm <- rbind(LCsumm,cbind(year = i, opens = init, closes = dissolve)) 
}

LCsumm <- as.data.frame(LCsumm)
print(LCsumm)
```

## Visualize the Data
The following creates an offline HTML page with the data.

```{r plotit, message=FALSE}
library(plotly)
p <- plot_ly(LCsumm, x = ~year, y = ~opens, type='bar', name= "Openings") %>%
        add_trace(y=~closes, name="Closings") %>%
        layout(title = "Larimer County Business Opens/Closes", yaxis = list(title='Count'), barmode= 'group')
print(p)
```





